Online (Bandit) Policy Selection for EASY-Backfilling

### 2017-03-31 paper/src/paper.tex
changes to eric's section.

\documentclass[sigconf]{acmart}
\setcopyright{rightsretained}

@@

Providing the computing infrastuctures needed to solve actual complex problems
arising in the various fields of the modern society (including climate change,
health, green energy or security) is a strategic challenge.  The main pillar to
address this challenge is to build extreme-scale parallel and distributed
platforms.  The race for always more computing power and storage capacity does
not only lead to sophisticated specific exascale platforms, but the objective
of the community is to design efficient sustained Petascale platforms.
However, there is still a long way to go and many scientific and technical
problems to solve by the system management software in order to adapt to such
large-scale evolution and the increasing complexity in both architecture and
applications.  The existing job and resource management softwares allow to run
tens thousands jobs on hundreds thousands cores.  They are based on robust
policies, however, despite their positive properties of relative simplicity and
their guarantee of no job starvation, there is still place for improvements.
We propose here to study how to tune the ordering of the submission queues in
the frame of the classical EASY-BackFilling family of heuristics using two new
online learning approaches (an approach based on a simulator and an approach
using multi-armed bandit).

Most common work in the literature consider \textit{a posteriori} optimization
of a scheduling algorithm on a (actual) data set.  The increasing amount and
diversity of data generated by large scale platforms motivates the community to
study more sophisticated policies that could leverage these data.  During the
last few years, there was an explosion of the number of works at the interface
of HPC and BigData fileds, dealing with learning algorithms .  Recent
sophisticated approaches developed \textit{ad hoc} supervised machine learning
algorithms for reducing uncertainty in problem parameters (e.g. job execution
times, memory requirements, etc.) via prediction.  However, only few works
study the impact of such techniques on the performances of the resource
manager.

@@

Our main contribution is to investigate two original methods for learning a
good scheduling policy on a per-system basis.  Each method aims to choose
the best queue reordering policy in a fixed search set of semantically diverse
options as fast as possible.

The first method (full feedback) continuously uses simulation on data gathered
from the past and current system workload. In order to reflect the
uncertainties in the data, we also evaluate a "noisy" variant in which
simulations are imprecise.

@@

\begin{figure*}[ht]
  \centering
  \includegraphics[scale=0.6]{figures/variability.pdf}
  \caption{Variability in the weekly average waiting time in the KTH-SP2 trace (see Subsection~\ref{sub:traces}) for 7 different policies.}
  \label{fig:mosn}
\end{figure*}

@@

We further assume that there is a certain regularity among periods, \textit{i.e.} that the distributions underlying the jobs submitted do not radically differ between consecutive periods. This assumption is validated in the study presented in~\ref{???}. It further entails that the behavior of a policy on the previous periods reflects its behavior on the current one, so that the selection of a policy can be based on its behavior on previous periods. However, there may be some variability between different periods for certain cost metrics. This is illustrated in Figure~\ref{fig:mosn} that displays the average waiting time for a trace using weekly periods encompassing one year. As one can note, the average waiting time varies a lot from one week to the other, for all the 7 policies considered. This indicates that when when the cost metric is averaged over different periods, there is a tradeoff to find in between longer periods that would somehow limit the variability, and shorter ones that yield more values for the estimation. We will come back to this issue below.

The selection of a policy is of course reminiscent of reinforcement learning. It is important to note, however, that a pure reinforcement learning approach is difficult to develop in our context as \textcolor{orange}{???}. We rely here on simpler, yet we believe more effective, strategies to solve this problem. These strategies are applied online and rely on exact simulation, noisy simulation and $\epsilon$-greedy bandit exploration.

@@

The above cost directly corresponds to the cumulative waiting time of the policy over the preceding periods (more precisely to the cumulative simulated estimate of the waiting time of the policy over the preceding periods) so that the length of the period has no impact here.

@@

In order to simulate how the simulation strategy for selecting policies would work on non simulated traces, we randomly introduce noise in the estimate of the waiting time defined by Eq.~\ref{eq:cost-exact} by rescaling it, either down or up, for each job:
%
\begin{enumerate}
\item For each job $j$ submitted during $\Delta_t$: $\rho_j \sim U[0.85,1.15]$
\item Then:
%
\begin{equation}
\label{eq:cost-noisy}
w_{\mbox{noisy}}(t;P_i) = \sum_{j=1}^{l(\Delta_t)} \rho_j \, \mbox{Wait}_j^{P_i}
\end{equation}
%
\end{enumerate}
%
$\rho$ is uniformly sampled in the interval $[0.85,1.15]$, thus adding a +/-15\% noise on the waiting time estimated by the simulator. The overall cost is then defined in the same way as before, leading to:
%
\begin{equation}
\label{eq:tot-cost-noisy}
w(\rightarrow T;P_i) = \sum_{t=0}^{T-1} \lambda^{T-1-t} w_{\mbox{noisy}}(t;P_i)
\end{equation}
%
As before (see Eq.~\ref{eq:select-exact}), the policy that minimizes the above cost is selected for the period $\Delta_T$.

Instead of introducing noise in the output of the simulator when selecting the policy, one could have introduced it when generating the execution traces. This would however have prevented us from comparing the exact and noisy selection methods, the first one representing an upper bound on the second one. As we will see in Section~\ref{sec:experiments}, there is little difference between the two selection methods, which is an important result of our study.

Lastly, as before, the length of the period has no impact here as the cost metric corresponds to the cumulative simulated waiting time for each policy.

@@

The previous strategies estimate the cost of all policies over all previous periods; the best policy according to this cost is then selected for the current period, the costs of all policies being updated for the next period. This is interesting as one maintains a complete knowledge of all policies over time. However, this requires computing many estimates at each period (as many as there are policies), which can be time consuming even with lightweight simulators.

We thus explore here a more efficient strategy that dispenses with estimating the cost of all policies at a given time and only updates the cost of the policy that is currently being used. This strategy makes use of the $\epsilon$-greedy exploration method, standard in reinforcement learning \cite{Watkins:1989}, to tradeoff exploitation and exploration, and relies, as before, on past estimates of the cost to select the best policy in the exploitation mode.

Let $l(\Delta_t)$ denote this time the number of jobs \textit{finished} during the period $\Delta_t$. We define the cost of policy $P_i$ during the period $\Delta_t$ as:
%
\[
w_{\epsilon}(t;P_i) =
\left\{
\begin{array}{ll}
 \sum_{j=1}^{l(\Delta_t)} \textbf{Wait}_j^{P_i} & \mbox{if} \, P_i \, \mbox{used during} \, \Delta_t \nonumber\\
 0 & \mbox{otherwise}
 \end{array}
\right.
\]
%
As one can note, the actual waiting time is used here, so that one dispenses with the use of a simulator for efficiency considerations. This indeed leads to faster and more accurate estimate of the cumulative waiting time, however only for the policy that is used during $\Delta_t$. This also explains the difference between the definitions of $l(\Delta_t)$: in the previous selection strategies, one relies on a simulated estimate so that jobs need not be finished during $\Delta_t$, but only submitted; in contrast, to know their actual waiting time, jobs need be finished during $\Delta_t$.

@@

The selection of the policy to be used during the current time period $\Delta_T$ is then based on the standard $\epsilon$-greedy exploration strategy:
%
\begin{itemize}
\item With probability $(1-\epsilon)$, select the policy $P$ that minimizes the cost over previous time periods (exploitation mode):
%
\begin{equation}
\label{eq:select-greedy}
P = \mbox{argmin}_{P_i, 1 \le i \le 10} w(\rightarrow T;P_i)
\end{equation}
%
\item With probability $\epsilon$, select a policy $P_i, \, 1 \le i \le 10$, at random (exploration mode).
\end{itemize}

Several studies propose to update $\epsilon$ over time, the exploration being less important once the estimates for the different policies are reliable \cite{Tokic:2010}. This strategy has however not been beneficial in our case. We believe this is due to the stochastic nature of the traces we use, as explained in Section~\ref{sec:experiments}.

>>>>

\documentclass[sigconf,review]{acmart}
\setcopyright{rightsretained}

@@

Providing the computing infrastuctures needed to solve actual complex problems
arising in the various fields of modern society (including climate change,
health, green energy or security) is a strategic challenge. The main pillar to
address this challenge is to build extreme-scale parallel and distributed
platforms. The race for always more computing power and storage capacity does
not only lead to sophisticated specific exascale platforms, but the objective
of the community is to design efficient sustained Petascale platforms. This
large-scale evolution and the increasing complexity in both architecture and
applications create many scientific and technical problems. Accordingly, system
management software still has a long way to go. The existing job and resource
management softwares allow to run tens thousands jobs on hundreds thousands
cores. They are based on robust policies, however, despite their positive
properties of relative simplicity and their guarantee of no job starvation,
there is still room for improvement. We propose here to study how to tune the
ordering of the submission queues in the frame of the classical
EASY-BackFilling family of heuristics using two new online learning approaches
(an approach based on a simulator and an approach using multi-armed bandit).

Most common work in the literature consider \textit{a posteriori} optimization
of a scheduling algorithm on a (actual) data set.  The increasing amount and
diversity of data generated by large scale platforms motivate the community to
study more sophisticated policies that could leverage these data.  During the
last few years, there was an explosion of the number of works at the interface
of HPC and BigData fileds, dealing with learning algorithms .  Recent
sophisticated approaches developed \textit{ad hoc} supervised machine learning
algorithms for reducing uncertainty in problem parameters (e.g. job execution
times, memory requirements, etc.) via prediction.  However, only few works
study the impact of such techniques on the performances of the resource
manager.

@@

Our main contribution is to investigate two original methods for learning a
good scheduling policy on a per-system basis. We focus in this work on the
average waiting time of the jobs. Each method aims to choose the best queue
reordering policy in a fixed search set of semantically diverse options.

The first method (full feedback) continuously uses simulation on data gathered
from the past and current system workload. In order to reflect the
uncertainties in the data, we also evaluate a "noisy" variant in which
simulations are imprecise\footnote{This mitigates the fact that we are
using the same batch scheduling simulator to learn and validate our policies}.

@@

\begin{figure*}[ht]
  \centering
  \includegraphics[scale=0.6]{figures/variability.pdf}

  \caption{Variability in the weekly average waiting time in the KTH-SP2 trace
  (see Subsection~\ref{sub:traces}) for 7 different policies. The policy set is
  reduced as not to obstruct the figure.}

  \label{fig:mosn}
\end{figure*}

@@

We further assume that there is a certain regularity among periods,
\textit{i.e.} that the distributions underlying the jobs submitted do not
radically differ between consecutive periods. This assumption is validated in
the study presented in~\cite{jsspp17}. It further entails that the behavior of
a policy on the previous periods reflects its behavior on the current one, so
that the selection of a policy can be based on its behavior on previous
periods. However, there may be some variability between different periods for
certain cost metrics~\cite{feitelson2001metrics}. This is illustrated in
Figure~\ref{fig:mosn} which displays the average waiting time of various
policies for the KTH-SP2 trace (see~\ref{sub:traces} for a description of the
workloads) using weekly periods encompassing one year. As one can note, the
average waiting time varies a lot from one week to the other, for all the 7
policies considered. This indicates that when when the cost metric is averaged
over different periods, there is a tradeoff to find in between longer periods
that would somehow limit the variability, and shorter ones that yield more
values for the estimation. We will come back to this issue below.

The selection of a policy is of course reminiscent of reinforcement learning.
It is important to note, however, that a pure reinforcement learning approach
is difficult to develop in our context. Indeed, while we have outlined a
reduced action space, the state space to consider is infinite. This is not
prohibitive \textit{per se}, as modern methods~\ref{approximatereinforcement}
can bypass dimensionality issues via function approximation. However, these
methods rely on large amounts of data, and we are studying online (often called
\textit{on-policy} methods in reinforcement learning) methods. We rely here on
simpler, yet we believe more effective, strategies to solve this problem.
These strategies are applied online and rely on exact simulation, noisy
simulation and $\epsilon$-greedy bandit exploration.

@@

The above cost directly corresponds to the cumulative waiting time of the
policy over the preceding periods (more precisely to the cumulative simulated
estimate of the waiting time of the policy over the preceding periods) so that
the length of the period has little impact here. The only bias comes from the
boundary states of the simulation. Indeed, we run simulations from an empty
system and wait for the system to be empty when job submissions cease at the
end of the submission period.

@@

In order to simulate how the simulation strategy for selecting policies would
work on non simulated traces, we randomly introduce noise in the estimate of
the waiting time defined by Eq.~\ref{eq:cost-exact} by rescaling it, either
down or up, for each job:
%
\begin{equation} \label{eq:cost-noisy} w_{\mbox{noisy}}(t;P_i) =
\sum_{j=1}^{l(\Delta_t)} \rho_j \, \mbox{Wait}_j^{P_i} \end{equation}
%
where $\rho$ is uniformly sampled in the interval $[0.85,1.15]$, thus adding a
+/-15\% noise on the waiting time estimated by the simulator. The overall cost
is then defined in the same way as before, leading to:
%
\begin{equation} \label{eq:tot-cost-noisy} w(\rightarrow T;P_i) =
\sum_{t=0}^{T-1} \lambda^{T-1-t} w_{\mbox{noisy}}(t;P_i) \end{equation}
%
As before (see Eq.~\ref{eq:select-exact}), the policy that minimizes the above
cost is selected for the period $\Delta_T$.

Instead of introducing noise in the output of the simulator when selecting the
policy, one could have introduced it when collecting the execution traces. Our choice
to add the noise after the simulation step is motivated by simplicity. As we
will see in Section~\ref{sec:experiments}, there is little difference between
the two selection methods, which is an important result of our study.

Lastly, as before, the length of the period has little impact here as the cost
metric corresponds to the cumulative simulated waiting time for each policy.

@@

The previous strategies estimate the cost of all policies over all previous
periods; the best policy according to this cost is then selected for the
current period, the costs of all policies being updated for the next period.
This is interesting as one maintains a complete knowledge of all policies over
time. However, this requires computing many estimates at each period (as many
as there are policies), which can be time consuming or cumbersome even with
lightweight simulators.

We thus explore here a more efficient strategy that dispenses with estimating
the cost of all policies at a given time and only updates the cost of the
policy that is currently being used. This strategy makes use of the
$\epsilon$-greedy exploration method, standard in reinforcement learning
\cite{Watkins:1989} and bandit problems~\cite{Auer2002}, to tradeoff
exploitation and exploration, and relies, as before, on past estimates of the
cost to select the best policy in the exploitation mode.

Let $l(\Delta_t)$ denote this time the number of jobs \textit{finished} during
the period $\Delta_t$. We define the cost of policy $P_i$ during the period
$\Delta_t$ as:
%
\[ w_{\epsilon}(t;P_i) = \left\{ \begin{array}{ll} \sum_{j=1}^{l(\Delta_t)}
  \textbf{Wait}_j^{P_i} & \mbox{if} \, P_i \, \mbox{used during} \, \Delta_t
  \nonumber\\ 0 & \mbox{otherwise} \end{array} \right.  \]
%
As one can note, the actual waiting time is used here, so that one dispenses
with the use of a simulator for efficiency considerations. This leads to a
faster and easier estimate of the cumulative waiting time, however only for the
policy that is used during $\Delta_t$. Note that there is as previously a bias
due to boundary effects in the measurement method, as we do include jobs that
had been started before $\Delta_t$, as well as disregard jobs that were
submitted during $\Delta_t$ but finish later.

%This also explains the difference
%between the definitions of $l(\Delta_t)$: in the previous selection strategies,
%one relies on a simulated estimate so that jobs need not be finished during
%$\Delta_t$, but only submitted; in contrast, to know their actual waiting time,
%jobs need be finished during $\Delta_t$.

@@

The selection of the policy to be used during the current time period
$\Delta_T$ is then based on the standard $\epsilon$-greedy exploration
strategy:

%
\begin{itemize}
\item With probability $(1-\epsilon)$, select the policy $P$ that minimizes the cost over previous time periods (exploitation mode):
%
\begin{equation}
\label{eq:select-greedy}
P = \mbox{argmin}_{P_i, 1 \le i \le 10} w(\rightarrow T;P_i)
\end{equation}
%
\item With probability $\epsilon$, select a policy $P_i, \, 1 \le i \le 10$, at random (exploration mode).
\end{itemize}

Several studies propose to decay $\epsilon$ over time, the exploration being
less important once the estimates for the different policies are reliable
\cite{Tokic:2010}. This strategy has however not been beneficial in our case.
We believe this is due to the properties of the traces we use, as
explained in Section~\ref{sec:experiments}.
